Explorations to Optimize the Parareal Algorithm \\to Solve ODEs in Given Applications

### 2014-12-10 paper.tex
paper updates

To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $gamma$ explained in the below table.

@@

For the speaker problem, we designed a problem-specific method to parallelize in space and divide Pierce Hall into subsections such that each processor would receive a division of the building to compute. We wish to show strong scaling, in which the computation time of our parallel method will decrease as we increase the number of processors. We compare this version to the serial version we did for Homework 4.

bksim TODO GET DATA, AS REQUIRED FROM ABOVE, SHOW GRAPHS

@@

As we can see, we do see the effects of strong scaling

@@

On the other hand, parallelism by space already requires a very problem-specific
design. The strong scaling efficiency of parallelism by space is high and is
conceptually embarrassingly parallel and easier to create. However, the
limitations are in the requirement for the code to be designed for the
problem---how to divide by space so as to allow for the maximum number of even
divisions. The lack of generality is not as beautiful as the parareal algorithm
but may lend itself to faster speedups.

>>>>

To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $\gamma$ explained in the below table.

@@

To demonstrate parallelism by space, we solve the two-dimensional wave equation by partioning a two-dimensional space domain amongst multiple processors. Specifically, we look to numerically solve

\begin{equation*}
\frac{\partial^2 u}{\partial t^2} = \frac{\partial^2 u}{\partial x^2} + \frac{\partial^2 u}{\partial y^2}
\end{equation*}

on the two-dimensional domain $(x, y) \in [0,1]\times[0,1]$ for evolution in time. We partition the domain by space using a second order central difference approximation, assigning a rectangular portion of the two-dimensional domain to each of a grid of $(P_x, P_y)$ processors. This requires the passing of information along the edges of the processors to each other, as shown graphically in the figure below:

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.5\linewidth]{decomp.png}
\caption{Parallelization by space schematic.}
\label{fig:spaceparallel}
\end{center}
\end{figure}

Note that each processor receives information from only those adjacent to it; that is, the top left processor in the graphic does not communicate directly with the bottom right processor in the graphic. Rather, information is passed only along shared boundary points to propagate the wave equation. We run this code for two different grid sizes, $128\times 128$ and $256\times 256$, as well as in serial.

In serial, we have:

\begin{center}
\begin{tabular}{|c|c|} \hline
($N_x, N_y$) & Time (seconds)\\ \hline
(128, 128) & 0.987\\ \hline
(256, 256) & 7.889\\ \hline
\end{tabular}
\end{center}

In parallel, we have:

\begin{center}
\begin{tabular}{|c|c|c|c|c|} \hline
($N_x, N_y$) & $P_x$ & $P_y$ & Time (sec) \\ \hline
(128, 128) & 2 & 2 &  \\ \hline
(128, 128) & 4 & 2 & 2.31 \\ \hline
(128, 128) & 4 & 4 &  \\ \hline
(128, 128) & 8 & 8 &  \\ \hline
(256, 256) & 2 & 2 & 2.57 \\ \hline
(256, 256) & 4 & 2 & 11.32 \\ \hline
(256, 256) & 4 & 4 & 13.65 \\ \hline
(256, 256) & 8 & 8 & 17.44 \\ \hline
\end{tabular}
\end{center}

Note that as the workload assigned to each processing element stays constant and more processors are used to solve a larger problem, linear scaling is achieved. However, scaling is \textbf{not} achieved when we simply add more processing elements to try to solve the same-sized problem, as there is greater constant startup costs which do not necessarily improve the speed (or perhaps even decrease speed). Hence, this is an example of weak scaling. However, an advantage to this paradigm is that because each processor only communicates with its nearest neighbors, communication overhead is relatively constant regardless of how large problem size / number of processors used becomes. Therefore, such a paradigm should scale well to larger problem sizes or more fine grid spacings.

@@

As we can see, we do see the effects of WEAK scaling

@@

On the other hand, parallelism by space already requires a more problem-specific
design. Parallelism by space is intuitive, and conceptually, is embarrassingly parallel and easier to create. However, the
limitations are in the requirement for the code to be designed for the
problem---how to divide by space so as to allow for the maximum number of even
divisions. The lack of generality is not as beautiful as the parareal algorithm
but may lend itself to faster speedups.
